Ages of Planets and Stars Observed by Kepler

### 2024-04-05 paper/ms.tex
correct missing Santos KOIs

We choose to adopt the results of \citetalias{Santos_2019} and
\citetalias{Santos_2021} as our default rotation periods.
\citetalias{Santos_2021} provides a detailed comparison against
\citetalias{McQuillan_2014}; the brief summary is that the periods
agree for 99.0\% of the 31{,}038 period detections in common between
the two studies.  \citetalias{Santos_2021} classified the 2{,}992
remaining stars from \citetalias{McQuillan_2014} as ``non-periodic''
based on updated knowledge of contaminants (e.g.~giant stars and
eclipsing binaries) and visual inspection.  In addition,
\citetalias{Santos_2021} report rotation periods for 24{,}182
main-sequence and subgiant FGKM stars that were not reported as
periodic by \citetalias{McQuillan_2014}.  Our speculation is that
these reported detections are more challenging than those from
\citetalias{McQuillan_2014}, perhaps due to a lower variability
amplitude, or a longer rotation period.  A relevant flag for stars
missing \citetalias{McQuillan_2014} periods is included in
Table~\ref{tab:stars}.
%TODO is it?
We verified that the main conclusions of our analysis do not change if
we were to adopt only the periods from \citetalias{McQuillan_2014}.
%TODO have you?

Concatenating the results of \citetalias{Santos_2019} and
\citetalias{Santos_2021} yields \nuniqstarsantosrot\ Kepler stars with
rotation periods.

@@

\begin{figure*}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.9\textwidth]{hist_samples_koi_gyro_ages_hist_field_gyro_ages_20230529_maxage3200.pdf}
	\end{center}
	\vspace{-0.6cm}
	\caption{
		{\bf Age distribution of stars and planet-hosts in the Kepler field
    inferred from rotation periods}.
    A histogram of samples from the age posteriors for all Kepler
    stars with reported rotation periods and temperatures is shown in
    light gray;
    the age calculation was actually valid for stars that meet the criteria
    discussed in Section~\ref{subsec:filters} (dark blue).
		\label{fig:hist_tgyro}
	}
\end{figure*}

@@

\startlongtable
\begin{deluxetable}{lllrrrr}
	%\tabletypesize{\footnotesize}
	\tabletypesize{\scriptsize}
	\tablecaption{Young planets and planet candidates.
		%The non-truncated machine readable versions are accessible both
		%through the online journal, and at
		%\url{https://zenodo.org/record/8327508}.
		\label{tab:youngplanets}}
	%\toprule
	%\midrule
	%\endhead
	\startdata
	KOI & Kepler & $t_{\rm gyro}$ & $R_{\rm p}$ & $P$ & $f_{\rm RUWE}$ & $f_{\rm grazing}$ \\
	-- &   -- & Myr &    Earths &    days &      bool &   bool  \\
	\hline
	\input{table_subgyr_confirmed} %\\
	%\hline
	\input{table_subgyr_candidate}
	\enddata
	\tablecomments{This table includes X, Y, Z... The machine-readable version,
		available online, includes additional columns for .... 
	}
\end{deluxetable}
%\endlongtable

%\clearpage

>>>>

We choose to adopt the results of \citetalias{Santos_2019} and
\citetalias{Santos_2021} as our default rotation periods.
\citetalias{Santos_2021} provides a detailed comparison against
\citetalias{McQuillan_2014}; the brief summary is that the periods
agree for 99.0\% of the 31{,}038 period detections in common between
the two studies.  \citetalias{Santos_2021} classified the 2{,}992
remaining stars from \citetalias{McQuillan_2014} as not showing
rotation periods based on updated knowledge of contaminants
(e.g.~giant stars and eclipsing binaries) and visual inspection.
In addition,
\citetalias{Santos_2021} report rotation periods for 24{,}182
main-sequence and subgiant FGKM stars that were not reported as
periodic by \citetalias{McQuillan_2014}.  Many of
these reported detections have lower variability amplitudes, and
longer periods, than those reported by
\citetalias{McQuillan_2014}.  A relevant flag for stars
missing \citetalias{McQuillan_2014} periods is included in
Table~\ref{tab:stars}.  Reviewing the resulting period detections, we
learned that 32 KOIs with orbital and rotation periods within
$\approx$20\% were excluded from the \citetalias{Santos_2019} and
\citetalias{Santos_2021} catalogs (A.~Santos, priv.~comm.) ; we choose
to include these objects.\footnote{In general, excluding stars in this
category would impose a bias against young close-in planets, due to
the natural commensurability between the rotation rates of young
stars, and Kepler's increased sensitivity to short-period objects.
While some of these KOIs are known false positives, our analysis
naturally excludes such objects at later steps.  As one motivating
example, this subset of stars happens to include {\it
e.g.}~Kepler~1643 (KIC~8653134), with $P_{\rm orb}$=5.34\,days and
$P_{\rm rot}$=5.05\,days, which is known to be $\approx$40\,Myr old
based on membership in RSG-5 \citep{Bouma_2022}.  }
The final tally of stars from \citetalias{Santos_2019},
\citetalias{Santos_2021}, and our corrected KOI list is
\nuniqstarsantosrot\ Kepler stars with rotation periods.

%  %TODO is it?
%  We verified that the main conclusions of our analysis do not change if
%  we were to adopt only the periods from \citetalias{McQuillan_2014}.
%  %TODO have you?

@@

\begin{figure*}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.9\textwidth]{hist_samples_koi_gyro_ages_hist_field_gyro_ages_20240405_maxage3200.pdf}
	\end{center}
	\vspace{-0.6cm}
	\caption{
		{\bf Age distribution of stars and planet-hosts in the Kepler field
    inferred from rotation periods}.
    A histogram of samples from the age posteriors for all Kepler
    stars with reported rotation periods and temperatures is shown in
    light gray;
    the age calculation was actually valid for stars that meet the criteria
    discussed in Section~\ref{subsec:filters} (dark blue).
		\label{fig:hist_tgyro}
	}
\end{figure*}

@@

\clearpage

\startlongtable
\begin{deluxetable}{lllrrrr}
	%\tabletypesize{\footnotesize}
	\tabletypesize{\scriptsize}
	\tablecaption{Young planets and planet candidates.
		%The non-truncated machine readable versions are accessible both
		%through the online journal, and at
		%\url{https://zenodo.org/record/8327508}.
		\label{tab:youngplanets}}
	%\toprule
	%\midrule
	%\endhead
	\startdata
	KOI & Kepler & $t_{\rm gyro}$ & $R_{\rm p}$ & $P_{\rm orb}$ & $f_{\rm RUWE}$ & $f_{\rm grazing}$ \\
	-- &   -- & Myr &    Earths &    days &      bool &   bool  \\
	\hline
	\input{table_subgyr_confirmed} %\\
	%\hline
	\input{table_subgyr_candidate}
	\enddata
	\tablecomments{This table includes X, Y, Z... The machine-readable version,
		available online, includes additional columns for .... 
	}
\end{deluxetable}
%\endlongtable

\clearpage
